### 2015-05-12 whoop.tex
fixes

To achieve this, we first \emph{instrument} a thread template for lockset analysis.  Given an arbitrary symbol, $i$, we define the instrumentation of a template $T$ with respect to $i$, denoted $\instrument{T}{i}$.  There are two aspects to this instrumentation phase: \emph{renaming} and \emph{lockset instrumentation}.

@@

Lockset instrumentation first requires the introduction of sets to track those variables that have been read from and written to, the mutexes that have been consistently used to protect each variable, and the set of current locks held by a thread.  To this end, when instrumenting template $T$ with respect to symbol $i$, we add sets $\hasbeenread{i} \subseteq \powerset{\sharedvars}$ and $\hasbeenwritten{i} \subseteq \powerset{\sharedvars}$ to track those shared variables that have been read from and written to, respectively, by the thread executing $T$.  We add a current lock set, $\currentlockset{i} \subseteq \powerset{\mutexes}$, to record the mutexes that are currently held by the thread, and for each shared variable $s \in \sharedvars$ we add a lock set, $\lockset{s}{i}$, to record the mutexes that are consistently held when the thread accesses $s$.

The statements of each procedure in $\instrument{T}{i}$ that manipulate shared variables and mutexes are then instrumented to take account of lock sets.  This transformation is described in Figure~\ref{fig:instrumentation}, where for an expression $e$ we use $\instrument{e}{i}$ to denote $e$ after renaming with respect to $i$.  A shared variable assignment, $s = e$, is instrumented by recording in $\hasbeenwritten{i}$ that $s$ has been written to, and updating the lock set for $s$, $\lockset{s}{i}$, to eliminate any mutexes that are not currently held (i.e.\ those mutexes that are not in $\currentlockset{i}$).  Notice that in the instrumented form, the shared variable $s$ is no longer assigned to.  We discuss the reasons for this when we describe \emph{shared state abstraction} below.  A shared variable read, $x = s$ operates analogously, except for the additional $\havockeyword$ command applied to the receiving private variable $x$; again this relates to shared state abstraction and is discussed below.  Instrumentation of mutex manipulation commands, $\mutexlock{m}$ and $\mutexunlock{m}$, involves updating the current lock set, $\currentlockset{i}$, to add and remove mutex $m$, respectively.

\medskip\noindent\textbf{Shared state abstraction }
%
Recall that while our aim is to perform race analysis for pairs of threads, we must be sure to account
for possible side-effects due to other threads that are running concurrently.  The instrumentation
of Figure~\ref{fig:instrumentation} achieves this via \emph{nondeterminism}: when reading from a shared variable $s$,
a nondeterministic value is returned.  This is reflected by the use of a $\havockeyword$ command,
which sets its argument to an arbitrary value.  Because all shared state accesses are abstracted in
this fashion, it is possible to completely dispense with the shared variables after instrumentation
has been performed.  As a result, when instrumenting a shared variable write, the effect of the write
is not explicitly modeled.

>>>>

To achieve this, we first \emph{instrument} a thread template for lockset analysis (see Section~\ref{bg:lockset}).  Given an arbitrary symbol, $i$, we define the instrumentation of a template $T$ with respect to $i$, denoted $\instrument{T}{i}$.  There are two aspects to this instrumentation phase: \emph{renaming} and \emph{lockset instrumentation}.

@@

Initially, lockset instrumentation introduces sets that track: the shared variables that have been read from and written to; the mutexes that have been consistently used to protect each variable; and the mutexes that are currently held by a thread.  To this end, when instrumenting template $T$ with respect to symbol $i$, we add the sets $\hasbeenread{i} \subseteq \powerset{\sharedvars}$ and $\hasbeenwritten{i} \subseteq \powerset{\sharedvars}$ to track the shared variables that have been read from and written to, respectively, by the thread executing $T$. For each shared variable $s \in \sharedvars$ we add a lockset, $\lockset{s}{i}$, to record the mutexes that are consistently held when the thread accesses $s$. Finally, we add a current lockset, $\currentlockset{i} \subseteq \powerset{\mutexes}$, to record the mutexes that are currently held by the thread.

The statements of each procedure in $\instrument{T}{i}$ that manipulate shared variables and mutexes are then instrumented to take account of the above sets.  This transformation is described in Figure~\ref{fig:instrumentation}, where for an expression $e$ we use $\instrument{e}{i}$ to denote $e$ after renaming with respect to $i$.  A shared variable assignment, $s = e$, is instrumented by recording in $\hasbeenwritten{i}$ that $s$ has been written to, and updating the lockset for $s$, $\lockset{s}{i}$, to eliminate any mutexes that are not currently held (i.e.\ those mutexes that are not in $\currentlockset{i}$).  Notice that in the instrumented form, the shared variable $s$ is no longer assigned to.  We discuss the reasons for this when we describe \emph{shared state abstraction} below.  A shared variable read, $x = s$ operates analogously, except for the additional $\havockeyword$ command applied to the receiving private variable $x$; again this relates to shared state abstraction and is discussed below.  Instrumentation of mutex manipulation commands, $\mutexlock{m}$ and $\mutexunlock{m}$, involves updating the current lockset, $\currentlockset{i}$, to add and remove mutex $m$, respectively.

\medskip\noindent\textbf{Shared state abstraction }
%
Recall that while our aim is to perform race analysis for pairs of threads, we must be sure to account for possible side-effects due to other threads that are running concurrently.  The instrumentation of Figure~\ref{fig:instrumentation} achieves this via \emph{nondeterminism}: when reading from a shared variable $s$, a nondeterministic value is returned.  This is reflected by the use of a $\havockeyword$ command, which sets its argument to an arbitrary value.  Because all shared state accesses are abstracted in this fashion, it is possible to completely dispense with the shared variables after the lockset instrumentation has been performed.  As a result, when instrumenting a shared variable write, the effect of the write is not explicitly modeled.
